---
title: "Credit Risk Modeling"
author: "Koki Ando"
date: "4/17/2019"
output: html_document
---

[Prob Space](https://prob.space/c/credit_default_risk)
[kaggle page](https://www.kaggle.com/uciml/default-of-credit-card-clients-dataset/discussion)


```{r}
library(naniar)
library(tidyverse)
library(ranger)
library(randomForest)
library(caret)
library(xgboost)
library(tictoc)
```


```{r}
test = read.csv("test_data.csv")
train = read.csv("train_data.csv")
names(train)[c(1, 25)] = c("ID", "Y")
test$Y = NA
credit = rbind(train, test)
names(credit) = c("ID", "Amount", "Sex", "Education", "Married",  "Age", "Past9",  "Past8",  "Past7", "Past6",  "Past5", "Past4", "Bill9", "Bill8", "Bill7", "Bill6", "Bill5", "Bill4", "Pre9", "Pre8", "Pre7", "Pre6", "Pre5", "Pre4", "Y")
credit = credit %>% 
  mutate(Total_Bill = Bill9+Bill8+Bill7+Bill6+Bill5+Bill4,
         Total_Pre = Pre9+Pre8+ Pre7+Pre6 +Pre5+Pre4)
credit = credit %>% 
  mutate(Bill_changed9 = (Bill9-Bill8)/Bill8,
         Bill_changed8 = (Bill8-Bill7)/Bill7,
         Bill_changed7 = (Bill7-Bill6)/Bill6,
         Bill_changed6 = (Bill6-Bill5)/Bill5,
         Bill_changed5 = (Bill5-Bill4)/Bill4)
credit[is.na(credit$Bill_changed9), ]$Bill_changed9 = 0
credit[is.na(credit$Bill_changed8), ]$Bill_changed8 = 0
credit[is.na(credit$Bill_changed7), ]$Bill_changed7 = 0
credit[is.na(credit$Bill_changed6), ]$Bill_changed6 = 0
credit[is.na(credit$Bill_changed5), ]$Bill_changed5 = 0
credit[is.infinite(credit$Bill_changed9), ]$Bill_changed9 = 0
credit[is.infinite(credit$Bill_changed8), ]$Bill_changed8 = 0
credit[is.infinite(credit$Bill_changed7), ]$Bill_changed7 = 0
credit[is.infinite(credit$Bill_changed6), ]$Bill_changed6 = 0
credit[is.infinite(credit$Bill_changed5), ]$Bill_changed5 = 0

credit = credit %>% 
  mutate(change_ave = (Bill_changed9+Bill_changed8+Bill_changed7+Bill_changed6+Bill_changed5)/5) %>% 
  select(-contains("Bill_changed"))
credit = credit %>% 
  mutate(final_change = (Bill9-Bill8)/Bill8)
credit[is.na(credit$final_change),]$final_change = 0
credit[is.infinite(credit$final_change),]$final_change = 0
# final change
credit[which(credit$final_change == -33598),]$final_change = -5.254125
credit[which(credit$final_change == max(credit$final_change)),]$final_change = -5.254125
credit[which(credit$final_change == max(credit$final_change)),]$final_change = -5.254125
# ave change
credit[which(credit$change_ave == min(credit$change_ave)),]$change_ave = -5.457273
credit[which(credit$change_ave == min(credit$change_ave)),]$change_ave = -5.457273
credit = credit %>% 
  mutate(Pay9 = ifelse(Past9 == -2, 1, 0), 
         Pay8 = ifelse(Past8 == -2, 1, 0), 
         Pay7 = ifelse(Past7 == -2, 1, 0),
         Pay6 = ifelse(Past6 == -2, 1, 0),
         notpaid = Pay9+Pay8+Pay7+Pay6)
credit = credit %>% 
  mutate(Young_married = ifelse(Age < 26 & Married == 1, 1, 0))
'%!in%' <- function(x,y)!('%in%'(x,y))
credit = credit %>% 
  mutate(low_edu = ifelse(Education %in% c(3,4), 1, 0))
credit = credit %>% 
  mutate(high_edu = ifelse(Education == 1, 1, 0))
credit = credit %>% 
  mutate(retired = ifelse(Age >= 60, 1, 0))
credit = credit %>% 
  mutate(age_bin = cut(Age, breaks = seq(20, 80, by = 10)))
# Normalization
processed_var = preProcess(select(credit, Total_Bill), method = "YeoJohnson")
credit = predict(processed_var, credit)
processed_var = preProcess(select(credit, Total_Pre), method = "YeoJohnson")
credit = predict(processed_var, credit)

# Amount ----------------------------
credit = credit %>%
  mutate(Amount_bin = ntile(Amount, 5))
processed_var = preProcess(select(credit, Amount), method = "YeoJohnson")
credit = predict(processed_var, credit)
# ----------------------------------------------
credit = credit %>% 
  mutate(age_bin_5 = cut(Age, breaks = seq(20, 80, by = 5)))
credit = credit %>% 
  mutate(youg_low_edu = ifelse(Age < 28 & low_edu == 1, 1, 0))
credit[which(credit$Bill4 == -339603),]$Bill4 = 18129
credit[which(credit$Bill7 == 1664089),]$Bill7 = 20330
credit[which(credit$Bill4 == -94625),]$Bill4 = 18131.5
credit[which(credit$Bill5 == 927171),]$Bill5 = 18104.5
processed_var = preProcess(select(credit, contains("Bill"), -Total_Bill), method = "YeoJohnson")
credit = predict(processed_var, credit)
credit = credit %>% 
  select(-contains("Pay"))
credit = credit %>% 
  mutate(total_past4 = ifelse(Past4 > 0, Past4, 0),
         total_past5 = ifelse(Past5 > 0, Past5, 0),
         total_past6 = ifelse(Past6 > 0, Past6, 0),
         total_past7 = ifelse(Past7 > 0, Past7, 0),
         past_count4_to_7 = total_past4+total_past5+total_past6+total_past7) %>% 
  select(-contains("total_past"))
```

upsampling

ステップの検証
手法の論文検証
1のEDA (支払い不履行の人の特徴を見つけ出す。)

```{r fig.width = 12, fig.height=4}
credit %>% 
  select(-c(Sex, Education, Married, ID, Young_married, Amount_bin, retired, notpaid), -contains("Past"), -contains("edu")) %>% 
  select_if(is.numeric) %>% 
  select(Age, Amount, Y) %>% 
  gather(key, value, -Y) %>% 
  ggplot(aes(factor(Y), value, fill = factor(Y))) +
  geom_boxplot() +
  facet_wrap(~key, scales = "free")
```


```{r}
boxplot(credit$Amount)
max(credit$Amount)
median(credit$Amount)
credit[which(credit$Amount == max(credit$Amount)), ]$Amount 

credit %>% 
  filter(Y == 0) %>% 
  summarise(mean(Amount), median(Amount))
```


##### Feature selection

```{r fig.width=8, fig.height=6}
result = read.csv("accuracy_table.csv")
result$trial = names(credit)[c(2:37)]
result %>%
  na.omit() %>% 
  gather(key, value, -trial) %>% 
  ggplot() +
  geom_boxplot(aes(reorder(trial, -value), value, fill = trial)) +
  coord_flip()
result %>%
  na.omit() %>% 
  mutate(ave_accuracy = (Accuracy+Accuracy.1+Accuracy.2+Accuracy.3)/4) %>% 
  select(trial, ave_accuracy) %>% 
  arrange(ave_accuracy)
```

### important features

```{r}
credit[, c(2:6, 36:37)] %>% 
  head
```

amount, sex, education, married, age


### meaningless features
# gbm modeling

```{r}
tic()
train = credit[!is.na(credit$Y),]
train$Y = factor(train$Y)
fitControl <- trainControl(method = "repeatedcv", number = 3, repeats = 5)
gbm_model = train(Y ~ .,
                  data = select(train, -ID),
                  method = "gbm", 
                  trControl = fitControl, verbose = FALSE)
# plot(gbm_model)
train$pred = predict(gbm_model, select(train, -Y, -ID))
confusionMatrix(train$Y, train$pred)
toc()
```


# cv xgboost modeling and evaluation

```{r}
tic()
train = credit[!is.na(credit$Y),]
set.seed(123)
train_index = sample(train$ID, nrow(train)*.75)
cv_train = filter(train, ID %in% train_index)
cv_test = filter(train, ID %!in% train_index)

cv <- xgb.cv(data = as.matrix(select(cv_train, -ID, -Y)), 
            label = cv_train$Y,
            nrounds = 100,
            nfold = 5,
            objective = "reg:linear",
            eta = .3,
            max_depth = 6,
            early_stopping_rounds = 10,
            verbose = 0
)
model_xgb <- xgboost(data = as.matrix(select(cv_train, -ID, -Y)), 
                     label = cv_train$Y,
                     nrounds = cv$best_iteration,
                      objective = "reg:linear",
                      eta = .3,
                      depth = 6
                     )
cv_test$pred = predict(model_xgb, as.matrix(select(cv_test, -ID, -Y)))
cv_test$pred = ifelse(cv_test$pred > .5, 1, 0)
confusionMatrix(as.factor(cv_test$Y), as.factor(cv_test$pred))$overall[1]
cv_test %>%
  mutate(error = pred - Y) %>%
  summarise(xgb_rmse = sqrt(mean(error^2)))
toc()
```

2: 0.8195556, 0.4247875
2: 0.8228148, 0.4209337 (9.852 sec elapsed)

# xgboost model

```{r}
credit_dummy = dummyVars(~., data = credit)
credit = as.data.frame(predict(credit_dummy, credit))
test = credit[is.na(credit$Y),] %>% 
  select(-Y)
train = credit[!is.na(credit$Y),]
cv <- xgb.cv(data = as.matrix(select(train, -ID, -Y)), 
            label = train$Y,
            nrounds = 100,
            nfold = 5,
            objective = "reg:linear",
            eta = .3,
            max_depth = 6,
            early_stopping_rounds = 10,
            verbose = 0
)
model_xgb <- xgboost(data = as.matrix(select(train, -ID, -Y)), 
                     label = train$Y,
                     nrounds = cv$best_iteration,
                      objective = "reg:linear",
                      eta = .3,
                      depth = 6,
                      verbose = 0
)
train$pred = predict(model_xgb, as.matrix(select(train, -Y, -ID)))
train$pred = ifelse(train$pred > .5, 1, 0)
confusionMatrix(as.factor(train$Y), as.factor(train$pred))$overall[1]
train %>%
  mutate(error = pred - Y) %>%
  summarise(xgb_rmse = sqrt(mean(error^2)))
```

1: 0.8397037, 0.4003702 [0.827] 117/283

## xgboost submission

```{r}
test$Y = predict(model_xgb, as.matrix(select(test, -ID)))
test$Y = ifelse(test$Y > .5, 1, 0)
submission = select(test, ID, Y)
write.csv(submission, "submission.csv", row.names = F)
```




# Random forest model

```{r}
tic()
test = credit[is.na(credit$Y),] %>% 
  select(-Y)
train = credit[!is.na(credit$Y),]
rf_model = randomForest(Y~., data = select(train, -ID))
train$pred = predict(rf_model, select(train, -ID, -Y))
train$pred = ifelse(train$pred > .5, 1, 0)
confusionMatrix(as.factor(train$Y), as.factor(train$pred))$overall[1]
toc()
```

1: 0.9902963, 0.09850738 [0.827] 119/282
2: 0.9905556, 0.09718253
3: 0.9948889, 0.07149204 [0.827] 117/282
4: 0.9955926 
5: 0.9962222 (681.673 sec elapsed)
6: 0.996963  (1891.157 sec elapsed) [0.829] 105/285
7: 

## Hyperparameter Tuning using caret

```{r}
tic()
test = credit[is.na(credit$Y),] %>% 
  select(-Y)
train = credit[!is.na(credit$Y),]
fitControl = trainControl(method = "repeatedcv", 
                          number = 3, repeats = 5)
rndomForest_model = train(
  Y ~ .,
  data = select(train, -ID),
  method = "rf", trControl = fitControl, verbose = FALSE
)
train$pred = predict(rndomForest_model, select(train, -ID, -Y))
train$pred = ifelse(train$pred > .5, 1, 0)
confusionMatrix(as.factor(train$Y), as.factor(train$pred))$overall[1]
toc()
```


## random forest submission

```{r}
test$Y = predict(rf_model, select(test, -ID))
test$Y = ifelse(test$Y > .5, 1, 0)
submission = select(test, ID, Y)
write.csv(submission, "submission.csv", row.names = F)
```

# ranger model

```{r}
tic()
test = credit[is.na(credit$Y),] %>% 
  select(-Y)
train = credit[!is.na(credit$Y),]
ranger_model = ranger(Y ~ .,
                      data=select(train, -ID), num.trees = 500,
                      respect.unordered.factors = "order")
train$pred = predict(ranger_model, select(train, -ID, -Y))$predictions
train$pred = ifelse(train$pred > .5, 1, 0)
confusionMatrix(as.factor(train$Y), as.factor(train$pred))$overall[1]
# train %>% 
#   mutate(error = pred-Y) %>% 
#   summarise(RMSE = sqrt(mean(error^2)))
toc()
```

1: 0.9902963, 0.09850738 [0.827] 119/282
2: 0.9905556, 0.09718253
3: 0.9948889, 0.07149204 [0.827] 117/282
6: 0.9914444 (44.135 sec elapsed)
7: 0.9918148
8: 0.9922593 
9: 0.9926667
10: 0.9928889
11: 0.9927407
12: 0.9932593 
13: 0.9938148 [0.829] 105/285 (46.617 sec elapsed)
14: 0.9935556 

## ragner submission

```{r}
test$Y = predict(ranger_model, select(test, -ID))$predictions
test$Y = ifelse(test$Y > .5, 1, 0)
submission = select(test, ID, Y)
write.csv(submission, "submission.csv", row.names = F)
```





# Model Stacking

0.996963, 0.829 randomForest
0.8382222, 0.826 xgboost
0.9935556, 0.822 ranger


